Typed Interactions for NLU in Opal

### 2017-12-13 paper.tex
More work on various sections

\subsection{Types for NLU}
If one were to try to implement our reminder interface na\"ively, a problem
would arise: the popular backends provide untyped interfaces and require
external configuration. Since Opal hopes to provide a consistent and safe
interface to multiple NLU engines, we decided to develop a typed abstraction for
natural language systems.

In this work, we present a domain specific language (DSL) for simultaneously
specifying
\begin{itemize}
\item configuration for an NLU engine;
\item a typed interface for responses
\end{itemize}
We also implement translations from our DSL to various artifacts which can be
integrated with Opal applications.
\\

@@

\begin{figure}
\begin{minted}{typescript}
  free-text Msg;
  keywords Time = "later" | "tomorrow";
  alias Set = {
    msg: Msg,
    time: Time
  };
  trait Intent =
    | <Set> Set
    | <Get> {};
\end{minted}
  \caption{Some DSL declarations.}
  \label{fig:dsl_decls}
\end{figure}

@@

\begin{figure}
\begin{minted}{typescript}
  {
    Intent: "Set",
    Msg: "pick up the dry cleaning",
    Time: "later"
  }
\end{minted}
  \caption{The unstructured response for the utterance: \emph{Remind me later to
      pick up the dry cleaning}.}
  \label{fig:nostructure}
\end{figure}

In order to reconstruct the desired data from the NLU backend's response, we
realized that it was necessary to keep data from the DSL around at run-time, not
just during setup.

\begin{figure*}
  \centering
  \begin{tikzpicture}[node distance = 3cm, auto]
    \node [block] (types) {Type DSL};
    \node [block, right of=types] (decls_hs) {Decl (hs)};
    \node [block, right of=decls_hs] (decls_ts) {Runtypes (ts)};
    \node [block, below of=decls_ts] (cfg) {NLU Config};
    \node [block, above of=decls_ts] (ts_types) {Types (ts)};
    \node [cloud, right of=decls_ts] (ts) {Opal};
    \node [cloud, right of=cfg] (backend) {Wit};

    \path [line] (types) -- (decls_hs);
    \path [line] (decls_hs) -- (decls_ts);
    \path [line] (decls_hs) -- (cfg);
    \path [line] (decls_hs) -- (ts_types);
    \path [line] (decls_ts) -- (ts);
    \path [line] (ts_types) -- (ts);
    \path [line] (cfg) -- (backend);
  \end{tikzpicture}
  \caption{The configuration process.}
  \label{fig:process}
\end{figure*}

We elected to use a TypeScript library called \emph{runtypes} to serialize the
AST directly into TypeScript; this gives us a data structure to guide response
parsing, and also provides us a way to dynamically type-check the results. We
discuss the response parsing algorithm in more detail in Section
\ref{implementation}.

Figure \ref{fig:process} shows process of configuring an application for NLU
interaction. The programmer writes types in our DSL, and uses our Haskell tool
to parse the declarations. The tool serializes three artifacts: a set of
TypeScript type declarations, a runtypes datastructure, and a set of JSON
configuration files for Wit. The programmer simply imports the types and
runtypes into their Opal project, uploads the JSON files to Wit, and the system
is ready to go.

\section{Formalism} \label{formalism}
In order to reason about the correctness of our translations, we have formalized
the results from Section \ref{description}. We begin by formalizing the type DSL
as shown in Figure \ref{fig:grammar_a}.
\begin{figure}
  \centering
  \begin{subfigure}{1\linewidth}
    \begin{align*}
      \fcy{D} ::=&\ \text{FreeText}(\etag{t}) \\
      |&\ \text{Keywords}(\etag{t},\ [\ell]) \\
      |&\ \text{Trait}(\etag{t},\ \{s: \fcy{T}\}) \\
      \\
      \fcy{T} ::=&\ \fcy{D} \mid \ell \mid \{s: \fcy{T}\}
    \end{align*}
    $$ \etag{t} \in \ff{Tag} \qquad \ell, s \in \textbf{string} $$
    \caption{Type DSL}
    \label{fig:grammar_a}
  \end{subfigure}\vspace{1cm}
  \begin{subfigure}{1\linewidth}
    $$ \tau ::=\ \textbf{string} \mid \ell \mid \{s: \tau\} \mid \langle \ell
    \rangle $$
    \caption{Relevant Opal Types}
    \label{fig:grammar_b}
  \end{subfigure}
  \caption{The core language.}
\end{figure}
There are similarities between these definitions and the ones in Figure
\ref{fig:ast}, but the core language has some slightly different goals. First,
we remove aliases and require that all type definitions are written ``in-line.''
Next, rather than split the definition by allowing tags within $\fcy{T}$, we
also require that entity definitions are inlined. This means that the data a
user expects from the NLU backend can be represented by a single tree of type
$\fcy{D}$. Importantly, none of these simplifications are fundamental---this
formulation can represent the same types as the actual system. This approach
simply gives up some usability to simplify reasoning.

In addition to a grammar for the core language itself, in Figure
\ref{fig:grammar_b} we define a grammar of types, $\tau$. We can think of $\tau$
as representing the final form of the data that a user hopes to obtain, with no
information about the entities that generated the data. They map almost directly
into types that are used by the underlying Opal system. We add a few types that
are noteworthy: the type {\bf string} is added to account for FreeText entities;
and $\langle {\ell} \rangle$ represents a disjunction of literals, and covers
the case of Keywords entities.

As mentioned above, user's view of their system can be formalized as a top-level
declaration generated by $\fcy{D}$. Since this declaration encompases both type
and configuration information, we would like a way to compute just the
\emph{type} of data that the user expects from the NLU engine. We define the
following mutually recursive functions, which traverse a declaration and
construct a single type in $\tau$.
\begin{align*}
  \ff{interp}\ &:\ \fcy{D} \to \tau \\
  \ff{interpTy}\ &:\ \fcy{T} \to \tau
\end{align*}

With the basic building blocks defined, we can move forward in formalizing our
system. \fxnote{Should I write out implementations?}

@@

The \ff{lookup} function always returns a string, but the string might mean any
number of things. The data might be a particular literal, in the case of a
keywords or trait entity, or an entire phrase in the case of a free-text entity.
We define configurations as functions that map a subset of entity tags to the
type of data expected for those entities.
$$ \ff{Cfg} \equiv \ff{Tag} \rightharpoonup \tau $$
Now, we call a response well-formed with respect to $\ff{cfg} : \ff{Cfg}$ if and
only if it returns appropriate data on \ff{lookup}. More formally, for some $r :
\ff{Resp}$,
\begin{align*}
  &r\ \text{well-formed}_{\ff{cfg}} \iff \\
  &\forall (\etag{t} : \textsf{Tag}).\ \ff{cfg}(\etag{t})\ \text{defined} \Rightarrow \ff{lookup}(\etag{t}, r) : \ff{cfg}(\etag{t})
\end{align*}

When constructing a member of \ff{Cfg}, we must take care that the mapping
respects entity types. Thus, if \etag{t} is a tag for a free-text entity,
$\ff{cfg}(\etag{t})$ is allowed to be any string. If, instead, \etag{t} is a tag
for keywords or trait entities, $\ff{cfg}(\etag{t})$ should return a union of
literals that represents the appropriate data.

@@

In the event that $r$ is not well-formed with respect to the configuration
generated by $d$, $\ff{parse}(d, r)$ is undefined. Otherwise, we define
\ff{parse} by case analysis on the declaration.
\begin{align*}
  \ff{parse}(\text{FreeText}(\etag{t}), r) = \ff{lookup}(\etag{t}, r)
\end{align*}
This case is fairly straightforward. Since the interpretation of a free-text
entity is just a {\bf string}, we can simply look up the value of the
appropriate entity from the response.
\begin{align*}
  \ff{parse}(\text{Keywords}(\etag{t}, [\ell_1, \dots, \ell_n]), r) =\\
  \ff{lookup}(\etag{t}, r)
\end{align*}
Perhaps surprisingly, we can use the exact same definition in the Keywords case
that we did in the FreeText case. We are required to return a value of type
$\langle {\ell_1 \mid \dots \mid \ell_n} \rangle$; as long as $r$ is
well-formed, it will have been configured to have the same list of literals. We
are guaranteed to get one of them back from \ff{lookup}.
\begin{align*}
  \ff{parse}(\text{Trait}(\etag{t}, rec), r) = \\
  \ff{parseTy}(rec[\ff{lookup}(\etag{t}, r)], r)
\end{align*}
The final case for \ff{parse} simply matches the value in the response to an
index into the record, and interprets that record entry using \ff{parseTy}.

@@

\subsection{Correctness}
We would like to bring both parts of the formalism together to show that our
system works the way we would like it to. The main property that we would like
to show is that \ff{parse} is correct.
\begin{theorem}[\ff{parse} Correct]
  Let $d$ be a declaration in $\fcy{D}$, and generate a configuration, \ff{cfg},
  from $d$. Then,
  \begin{flalign*}
    &\forall\ (r : \textsf{Resp}).\\
    &r\ \text{well-formed}_{\ff{cfg}} \Rightarrow \ff{parse}(d, r)\ \text{defined}
  \end{flalign*}
\end{theorem}
\begin{proof}
  By induction on $d$.
\end{proof}
Note that since \ff{parse} is a dependent function, it will already return a
value of the correct type. Thus, we only need to care that it is defined.

\section{Implementation} \label{implementation}

@@

Our configuration tool is written in Haskell, and is designed to be simple and
obviously correct. We use the Parsec library to parse the type DSL into the
intermediate representation mentioned in Figure \ref{fig:hs_ast}.

@@

@@

In section \ref{formalism}, we outline the parse function, which takes a
declaration and a response and generates an Opal object. The function that we
provide in the Opal library is similar, but there are two important differences.
First, the parse function takes some element of $\fcy{D}$, which it uses to
deduce the structure of the desired output. In the real application, the analog
of $\fcy{D}$ is the \hs{Decl} type in the Haskell code; this is not available to
the Opal application at run time. As mentioned above, we get around this by
using a modified version the runtypes library to keep this information around at
run-time. We extend runtypes to include a type for entities---which map to
declarations in $\fcy{D}$. Runtypes provides the infrastructure for non-entity
types in $\fcy{T}$.

@@

\section{Evaluation} \label{evaluation}
%TODO: Cover limitations of the design, possible use-cases, etc

@@

\subsubsection{More Types}
In addition to the three ``universal'' kinds of entities (free-text, keywords,
and traits), the Wit NLU engine provides more specialized ways of looking up
entities. Wit provides support for parsing numbers, dates, times, and other
specific forms of text. We elected to keep things simple at first and exclude
these built-in entities, but adding support for them seems like a very
reasonable next-step.

@@

\bibliography{paper}{}
\bibliographystyle{plain}

>>>>

\subsection{Types for NLU}
If a programmer simply wrote down types like the ones above, set up their NLU
engine, and started their application, there would be a problem. Since there is
no connection between the types and the NLU configuration, these types could
actually be unsound! If the programmer made a mistake in \emph{either} the types
\emph{or} the configuration, the application would be unusable and no compiler
could possibly know. Bugs like these are particularly insidious because they
might be mistaken as network errors or even just as an indicator of
under-training.

In this work, we set out to connect the types and the NLU configuration,
allowing the programmer to avoid these awful bugs. There are a few ways that one
might try to do this. First, the programmer could write down the NLU
configuration and extract the types from it. This is pretty much a non-starter,
since there is fundamentally less information in the NLU configuration than
there is in the type definitions. Instead, the programmer could write the types,
and extract the NLU information from those. This approach is more feasible, but
it would require some kinds of annotations in the types to capture
implementation-specific properties of the configuration.

We present a third option: a domain specific language (DSL) for simultaneously
specifying
\begin{itemize}
\item configuration for an NLU engine;
\item a typed interface for responses
\end{itemize}
This allows the user to tune the configuration as much as they need, to write
useful types, and to ensure that these two artifacts are always compatible.
\\

@@

\begin{figure}
\begin{minted}{typescript}
  free-text Msg
  keywords Time = "later" | "tomorrow"
  alias Set = {
    msg: Msg,
    time: Time
  }
  trait Intent =
    | <Set> Set
    | <Get> {}
\end{minted}
  \caption{Some DSL declarations.}
  \label{fig:dsl_decls}
\end{figure}

@@

\begin{figure}
\begin{minted}{typescript}
  {
    Intent: ["Set"],
    Msg: ["pick up the dry cleaning"],
    Time: ["later"]
  }
\end{minted}
  \caption{The unstructured response for the utterance: \emph{Remind me later to
      pick up the dry cleaning}.}
  \label{fig:nostructure}
\end{figure}

This makes it necessary to have a run-time component of the system that knows
how to reconstruct structured data from the unstructured response that comes
from the NLU engine. Building structured data from the response is a nontrivial
task. First, it requires that information about the relationships between
entities and data be available at run-time. This is mostly an engineering
problem, and our solution is outlined in Section \ref{implementation}.

The other difficulty is that the types might admit more than one valid response
object. To see how this works, imagine that we want to allow a user to set two
messages at once. We would change the \ts{Set} type to be:
\begin{minted}{typescript}
  type Set = {
    msg1: Msg,
    msg2: Msg,
    time: Time
  };
\end{minted}
Unfortunately, this kind of ambiguity makes it impossible to find a principled
way of parsing a response. Since the backends do not expose ordering
information, a response with two \ts{Msg} fields might parse to any of four
different objects. We say four because the reconstruction algorithm might
reasonably ignore one of the messages and just put the same one for \ts{msg1}
and \ts{msg2}. For this reason, we impose the restriction that declarations do
not reuse entities.

\section{Formalism} \label{formalism}
In order to reason about the correctness of our translations, we have formalized
the results from Section \ref{description}. We begin by formalizing the type DSL
as shown in Figure \ref{fig:grammar_a}.
\begin{figure}
  \centering
  \begin{subfigure}{1\linewidth}
    \begin{align*}
      \fcy{D} ::=&\ \text{FreeText}(\etag{t}) \\
      |&\ \text{Keywords}(\etag{t},\ [\ell]) \\
      |&\ \text{Trait}(\etag{t},\ \{s: \fcy{T}\}) \\
      \\
      \fcy{T} ::=&\ \fcy{D} \mid \ell \mid \{s: \fcy{T}\}
    \end{align*}
    $$ \etag{t} \in \ff{Tag} \qquad \ell, s \in \textbf{string} $$
    \caption{Type DSL}
    \label{fig:grammar_a}
  \end{subfigure}\vspace{1cm}
  \begin{subfigure}{1\linewidth}
    $$ \tau ::=\ \textbf{string} \mid \ell \mid \{s: \tau\} \mid \langle \tau
    \rangle $$
    \caption{Relevant Opal Types}
    \label{fig:grammar_b}
  \end{subfigure}
  \caption{The core language.}
\end{figure}
There are similarities between these definitions and the ones in Figure
\ref{fig:ast}, but the core language has some slightly different goals. First,
we remove aliases and require that all type definitions are written ``in-line.''
Next, rather than split the definition by allowing tags within $\fcy{T}$, we
also require that entity definitions are inlined. This means that the data a
user expects from the NLU backend can be represented by a single tree of type
$\fcy{D}$. Importantly, none of these simplifications are fundamental---this
formulation can represent the same types as the actual system. This approach
simply gives up some usability to simplify reasoning.

In addition to a grammar for the core language itself, in Figure
\ref{fig:grammar_b} we define a grammar of types, $\tau$. We can think of $\tau$
as representing the final form of the data that a user hopes to obtain, with no
information about the entities that generated the data. They map almost directly
into types that are used by the underlying Opal system. We add a few types that
are noteworthy: the type {\bf string} is added to account for FreeText entities;
and $\langle {\tau} \rangle$ represents an arbitrary union, and covers the cases
of Keywords and Trait entities.

As mentioned above, user's view of their system can be formalized as a top-level
declaration generated by $\fcy{D}$. Since this declaration encompases both type
and configuration information, we would like a way to compute just the
\emph{type} of data that the user expects from the NLU engine. We define the
following mutually recursive functions, which traverse a declaration and
construct a single type in $\tau$.
\begin{align*}
  \ff{interp}\ &:\ \fcy{D} \to \tau \\
  \ff{interpTy}\ &:\ \fcy{T} \to \tau
\end{align*}
We implement \ff{interp} and \ff{interpTy} in Figure \ref{fig:interp}.

\begin{figure}
  \centering
  \begin{flalign*}
    & \ff{interp}(\text{FreeText}(\etag{t})) = \textbf{string} \\
    & \ff{interp}(\text{Keywords}(\etag{t}, [\ell])) = \langle \ell \rangle \\
    & \ff{interp}(\text{Trait}(\etag{t'}, \{s : t\})) = \\
    & \qquad \left< \{\text{tag}: s,\ \text{data}: \ff{interpTy(t)}\} \right> \\
    \\
    & \ff{interpTy}(d) = \ff{interp(d)} \\
    & \ff{interpTy}(\ell) = \ell \\
    & \ff{interpTy}(\{s: t\}) = \{s: \ff{interpTy}(t)\}
  \end{flalign*}
  \caption{Type interpretation.}
  \label{fig:interp}
\end{figure}

One interesting thing to point out is that the record in the Trait declarations
becomes a union type. This makes sense, since a trait represents a property of
the utterance as a whole, and thus can decide the structure of the response.

With the basic building blocks defined, we can move forward in formalizing our
system.

@@

The \ff{lookup} function always returns a string, but the string might mean any
number of things. The data might be a particular literal, in the case of a
keywords or trait entity, or an entire phrase in the case of a free-text entity.
We define configurations as functions that map a subset of entity tags to the
type of data expected for those entities. The function \ff{cfg} takes a
declaration and produces a configuration for that declaration.
$$ \ff{cfg}\ :\ \fcy{D} \to \ff{Tag} \rightharpoonup \tau $$
We define \ff{cfg} in Figure \ref{fig:cfg}. The exact details of the else branch
of the Trait case are implementation specific, but the basic idea to map the
\ff{cfg} function over all nested declarations. This achieves the goal of
capturing every tag in the declaration.

\begin{figure}
  \centering
  \begin{flalign*}
    & \ff{cfg}(\text{FreeText}(\etag{t}), \etag{t}) = \textbf{string} \\
    & \ff{cfg}(\text{Keywords}(\etag{t}, [\ell]), \etag{t}) = \langle \ell \rangle \\
    & \ff{cfg}(\text{Trait}(\etag{t}', \{s : t\}), \etag{t}) = \\
    & \qquad \textbf{if}\ \etag{t} = \etag{t}'\ \textbf{then}\ \langle {s} \rangle\ \textbf{else}\ \dots
  \end{flalign*}
  \caption{Configuation generation.}
  \label{fig:cfg}
\end{figure}

Now, we call a response well-formed with respect to $d : \fcy{D}$ if and only if
it returns appropriate data on \ff{lookup}. More formally, for some $r :
\ff{Resp}$,
\begin{align*}
  &r\ \text{well-formed}_{d} \iff \\
  &\ \ \forall (\etag{t} : \textsf{Tag}).\ \ff{cfg}(d, \etag{t})\ \text{defined} \Rightarrow  \\
  &\qquad \ff{lookup}(\etag{t}, r) : \ff{cfg}(d, \etag{t})
\end{align*}
We will use well-formedness as our basic assumption that the application has
been configured correctly relative to the response.

@@

In the event that $r$ is not well-formed with respect to the configuration
generated by $d$, $\ff{parse}(d, r)$ is undefined. Otherwise, we define
\ff{parse} by case analysis on the declaration.
\begin{align*}
  \ff{parse}(\text{FreeText}(\etag{t}), r) = \ff{lookup}(\etag{t}, r)
\end{align*}
This case is fairly straightforward. Since the interpretation of a free-text
entity is just a {\bf string}, we can simply look up the value of the
appropriate entity from the response.
\begin{align*}
  \ff{parse}(\text{Keywords}(\etag{t}, [\ell_1, \dots, \ell_n]), r) =\\
  \ff{lookup}(\etag{t}, r)
\end{align*}
Perhaps surprisingly, we can use the exact same definition in the Keywords case
that we did in the FreeText case. We are required to return a value of type
$\langle {\ell_1 \mid \dots \mid \ell_n} \rangle$; as long as $r$ is
well-formed, it will have been configured to have the same list of literals. We
are guaranteed to get one of them back from \ff{lookup}.
\begin{align*}
  \ff{parse}(\text{Trait}(\etag{t}, rec), r) = \\
  \ff{parseTy}(rec[\ff{lookup}(\etag{t}, r)], r)
\end{align*}
The final case for \ff{parse} simply matches the value in the response to an
index into the record, and interprets that record entry using \ff{parseTy}. It
is important to note that since \ff{interp} returns a union here, we only need
to return a value that is in some branch of that union. In this case, we use the
\ff{lookup} value to decide which branch.

@@

\subsection{Correctness}
We would like to bring both parts of the formalism together to show that our
system works the way we would like it to. The main property that we would like
to show is that \ff{parse} is correct.
\begin{theorem}[\ff{parse} Correct]
  Let $d$ be a declaration in $\fcy{D}$. Then,
  \begin{flalign*}
    &\forall\ (r : \textsf{Resp}).\\
    &r\ \text{well-formed}_d \Rightarrow \ff{parse}(d, r)\ \text{defined}
  \end{flalign*}
\end{theorem}
\begin{proof}
  By induction on $d$.
\end{proof}
Note that since \ff{parse} is a dependent function, it will already return a
value of the correct type. Thus, we only need to care that it is defined.

\section{Implementation} \label{implementation}
\begin{figure*}
  \centering
  \begin{tikzpicture}[node distance = 3cm, auto]
    \node [block] (types) {Type DSL};
    \node [block, right of=types] (decls_hs) {Decl (hs)};
    \node [block, right of=decls_hs] (decls_ts) {Runtypes (ts)};
    \node [block, below of=decls_ts] (cfg) {NLU Config};
    \node [block, above of=decls_ts] (ts_types) {Types (ts)};
    \node [cloud, right of=decls_ts] (ts) {Opal};
    \node [cloud, right of=cfg] (backend) {Wit};

    \path [line] (types) -- (decls_hs);
    \path [line] (decls_hs) -- (decls_ts);
    \path [line] (decls_hs) -- (cfg);
    \path [line] (decls_hs) -- (ts_types);
    \path [line] (decls_ts) -- (ts);
    \path [line] (ts_types) -- (ts);
    \path [line] (cfg) -- (backend);
  \end{tikzpicture}
  \caption{The configuration process.}
  \label{fig:process}
\end{figure*}
A high-level workflow for our implementation is shown in Figure
\ref{fig:process}. The programmer writes types in our DSL, and uses our Haskell
tool to parse the declarations. The tool serializes three artifacts: a set of
TypeScript type declarations, a runtypes datastructure, and a set of JSON
configuration files for Wit. The programmer simply imports the types and
runtypes into their Opal project, uploads the JSON files to Wit, and the system
is ready to go.

@@

Our configuration tool is written in Haskell, and is designed to be simple and
obviously correct. We use the Parsec library to parse the type DSL into the
intermediate representation mentioned in Figure \ref{fig:hs_ast}.\cite{parsec}

@@

After parsing with Parsec, we apply versions of the \ff{interp} and \ff{cfg}
functions from Section \ref{formalism} to generate the appropriate
configuration.

@@

In section \ref{formalism}, we outline the parse function, which takes a
declaration and a response and generates an Opal object. The function that we
provide in the Opal library is similar, but there are two important differences.
First, the parse function takes some element of $\fcy{D}$, which it uses to
deduce the structure of the desired output. In the real application, the analog
of $\fcy{D}$ is the \hs{Decl} type in the Haskell code; this is not available to
the Opal application at run time.

We get around this by using a modified version the a TypeScript library called
Runtypes.\cite{runtypes} Runtypes allows us to to serialize declarations
directly into TypeScript and therefore to keep necessary information around at
run-time. We extend Runtypes to include a type for entities---which map to
declarations in $\fcy{D}$. Runtypes provides the infrastructure for non-entity
types in $\fcy{T}$.

@@

\section{Evaluation} \label{evaluation}
Working with the system to build some toy applications led us to a number of
important conclusions about its effectiveness. We discuss a number of these
conclusions here.

\subsection{Type Expressiveness}
The first important thing we found when starting to write types in our DSL is
that the standard three entity types are not quite expressive enough for real
applications. For example, when trying to implement a small chatbot for
scheduling calendar meetings, we wanted to write something like,
\begin{minted}{typescript}
  free-text Person
  ...
  trait Intent = <Schedule> {
    person: Person,
    date: Date,
    time: Time
  }
\end{minted}
but we found that we had no way to express a date or a time. Technically
speaking, it might have been possible to encode dates and times as free-text and
do a second parsing pass in the application code, but that solution is far from
elegant. We discuss one potential solution to this problem in Section
\ref{future>types}.

\subsection{Programmer Freedom}
When working around unsupported types, there is an obvious tendency to modify
the application types to better fit the capabilities of the system. It turns out
that this occurs in more cases than just using dates and times. For example,
consider another calendar example, similar to the one above.
\begin{minted}{typescript}
keywords WkDay = "Sunday" | ...
alias Abs = {
  day: WkDay
  person: Person
}
keywords RDay = "today" | "tomorrow"
alias Rel = {
  day: RDay
  person: Person
}
trait Intent =
  | <Absolute> Abs
  | <Relative> Rel
\end{minted}
Here, we've done away with the complications of dates and times, and we just
allow a user to specify a day of the week or a ``relative day.'' The interesting
observation here is that this is likely not the way that most TypeScript/Opal
developers would write this type. For example, one reasonable alternative might
express days as a top-level disjunction, and then make the type of intents a
normal record.
\begin{minted}{typescript}
  type Day =
    | { tag: "today" }
    | { tag: "tomorrow" }
    | { tag: "abs", day: WkDay }
  type Intent =
    { day: Day, person: Person }
\end{minted}
This makes sense, since \ts{Absolute} and \ts{Relative} are not actully
different \emph{intents}, per se. We can see that the system has led us to a
different type than we would normally write.

It turns out that this might actually be a good thing, if viewed from the
perspective of the NLU engine itself. While \ts{Absolute} and \ts{Relative} are
the same intent from a philosophical perspective, they might be parsed very
differently. The utterance
\begin{center}
  \emph{Schedule my Wednesday meeting with Chris.}
\end{center}
is very different in structure from
\begin{center}
  \emph{Set a meeting with James tomorrow.}
\end{center}
Thus, it might be the case that forcing the programmer to write slightly
different types will make it much easier for the NLU backend to find the correct
information.

Overall, there is likely some work to be done in terms of making our types more
idiomatic, but for now we err on the side of making life easier for the NLU
engine.

@@

\subsubsection{More Types} \label{future>types}
In addition to the three ``universal'' kinds of entities (free-text, keywords,
and traits), the Wit NLU engine provides more specialized ways of looking up
entities. Wit provides support for parsing numbers, dates, times, and other
specific forms of text. We elected to keep things simple at first and exclude
these built-in entities, but adding support for them seems like a very
reasonable next-step.

@@

% TODO: Fix bib formatting
\bibliography{paper}{}
\bibliographystyle{plain}
